Dismal Code: Studying the Evolution of Security Bugs

### 2013-06-26 docs/evol.tex
minor changes

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12,TJBD11,M06}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs~\cite{SZ12}, little attention has been paid to study them in relation to
software evolution~\cite{L96,LRWPT97,IB06,RGMA06}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse how different
types of security vulnerabilities are related to the evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,HP04,HW08,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://mvnrepository.com/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

@@

In the rest of this paper we
describe the processing of our data and our experiment (Section \ref{sec:meth}),
present and discuss the results we obtained (Section \ref{sec:res}),
outline related work (Section \ref{sec:rel}),
and end up with a conclusion and directions for future work (Section \ref{sec:con}).

@@

\begin{table*}[t]
\centering
\caption{Number of Project Releases that Contain at Least One Severe Security Bug}
\label{tbl:sev}
\leavevmode
	\begin{tabular}{p{24em} r}
	\hline
	Bug Description & Number of Project Releases\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 151\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,579\\
	{\sc pt}: absolute path traversal in servlet  & 103\\
	{\sc pt}: relative path traversal in servlet & 57\\
	{\sc sql}: non-constant string passed to execute method on an {\sc sql} statement & 1,875\\
	{\sc sql}: a prepared statement is generated from a non-constant String & 1,486\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 18\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 90\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 142\\
	\hline
	\end{tabular}
\end{table*}

@@

\section{Acknowledgments}
This research has been co-financed by the European Union (European Social Fund
--– {\sc esf}) and Greek national funds through the Operational Program
``Education and Lifelong Learning'' of the National Strategic Reference Framework –
Research Funding Program: Heracleitus II. Investing in knowledge society
through the {\sc esf}.

>>>>

A security bug is a programming error that introduces a potentially exploitable
weakness into a computer system~\cite{SSL12,TJBD11,M06}. This
weakness could lead to a security breach with unfortunate consequences in
different layers, like databases, native code, applications, libraries and
others. Despite the significant effort to detect and eliminate such
bugs~\cite{SZ12}, little attention has been paid to study them in relation to
software evolution~\cite{L96,LRWPT97}. 
One of the most common approaches to identify security bugs is
{\it static analysis}~\cite{CW07}. This kind of analysis involves the
inspection of the program's source or object code without executing
it.

In this
paper we present how we used a large software ecosystem to analyse how different
types of security vulnerabilities are related to the evolving software packages.
For our research we used {\it FindBugs},\footnote{\url{http://findbugs.sourceforge.net/}}
a static analysis tool that examines bytecode to detect software bugs and has already been used in
research~\cite{AP10,HP07,HW08,SHP06}.
Specifically, we ran FindBugs on all the project
versions of all the projects that exist in the
{\it Maven Central Repository}\footnote{\url{http://mvnrepository.com/}}
(approximately 265GB of data---see Section~\ref{sec:data}).
Then we observed the changes that involved the security bugs and their characteristics.
This research builds upon our earlier work on the topic~\cite{MGS12}.

@@

In the rest of this paper we
outline related work (Section \ref{sec:rel}),
describe the processing of our data and our experiment (Section \ref{sec:meth}),
present and discuss the results we obtained (Section \ref{sec:res}),
and end up with a conclusion and directions for future work (Section \ref{sec:con}).

@@

\begin{table*}[t]
\centering
\caption{Number of Project Releases that Contain at Least One Severe Security Bug}
\label{tbl:sev}
\leavevmode
	\begin{tabular}{p{31em} r}
	\hline
	Bug Description & Number of Project Releases\\
 	\hline
	{\sc hrs}: {\sc http} cookie formed from untrusted input & 151\\
	{\sc hrs}: {\sc http} response splitting vulnerability & 1,579\\
	{\sc pt}: absolute path traversal in servlet  & 103\\
	{\sc pt}: relative path traversal in servlet & 57\\
	{\sc sql}: non-constant string passed to execute method on an {\sc sql} statement & 1,875\\
	{\sc sql}: a prepared statement is generated from a non-constant String & 1,486\\
	{\sc xss}: {\sc jsp} reflected cross site scripting vulnerability & 18\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability in error page & 90\\
	{\sc xss}: Servlet reflected cross site scripting vulnerability & 142\\
	\hline
	\end{tabular}
\end{table*}

@@

\section{Acknowledgments}
This research has been co-financed by the European Union (European Social Fund
- {\sc esf}) and Greek national funds through the Operational Program
``Education and Lifelong Learning'' of the National Strategic Reference Framework -
Research Funding Program: Heracleitus II. Investing in knowledge society
through the {\sc esf}.
